More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-06-26 2016-Weak-FH-Bisim/WeakBisim.tex
proof of weak bisim

INtiuition: tau is non observable if the automaton allows any tau transition from holes, and the global transition resulting from a tau in a hole is a tau transition not changing the action state.
\TODO{Id(s) is the identity function on variables of state s}
\begin{definition}[Non-observability of silent actions]
An open automaton $A = <J,\mathcal{S},s_0,\mathcal{T}>$ \emph{cannot observe silent actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\symb{True},\symb{Id}(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $\alpha$, $j$, $s$, $\Post$ and $s'$ such that
\[ \openrule
         {
           (j\mapsto\tau),\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] we have \[ \alpha=\tau\land s=s'\land \Post=\symb{Id}(s)
\]
\end{enumerate}
\end{definition}

@@

\begin{property}[Weak open transition transitivity]
Let $A$ be an open automaton, and $<J,\mathcal{S},s_0,\WT>$ a weak open automaton derived from $A$.
Suppose we have the following weak open transitions in $\WT$:
\[\openrule
         {
           \overline{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1}
\qquad 
\openrule
         {
           \overline{\gamma_2},\Pred_2,\Post_2   }
         {s_1 \OTWeakarrow {\alpha} s_2}
\qquad
\openrule
         {
           \overline{\gamma_3},\Pred_3,\Post_3   }
         {s_2 \OTWeakarrow {\tau} s_3}
\]
Then there is ag weak  open transition of the following form in $\WT$:
\[\openrule
         {
           \overline{\gamma},\Pred,\Post   }
         {s \OTWeakarrow {\alpha} s_3}
\]
Suppose additionally $\Post_3=({x_k\gets e_k})^{k\in K}$, we have the following
\[
\begin{array}{rcl}
\gamma&=& \gamma_1\dotcup\gamma_2\dotcup\gamma_3\\
\Pred&=&\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}\\
\Post &=& ({x_k\gets ((e_k\subst{\Post_2})\subst{\Post_1})})^{k\in K}
\end{array}\]
\end{property}

@@

Weak bisimulation can be alternatively defined as a strong bisimulation on the weak open automata, however this would still need a specific definition of strong bisimulation for weau open automata. \TODO{polish}

@@

\subsection{Running example}
- Detail of a $\tau^*$ sequence, \\
- construction of the predicate\\
- defn / elimination of the intermediaite variables\\
- construction of the weak OT.

The ``saturation'' step in the construction of the weak relation will
pick all $\tau^*$ and $\tau^*;a;\tau^*$ sequences from the OA, and build a weak
OT from each sequence. Here is one example of a short sequence,
starting with the open transition $OT_2$ of page \pageref{OT:ABP},
then transmitting the message to the receiver (if the sender and
receiver control bits are equal), and to the environnement. The two
firts OTs are internal, the last one is visible, so it constitute a
$\tau;\tau;a$ weak transition.

>>>>

Intuition: tau is non observable if the automaton allows any tau transition from holes, and the global transition resulting from a tau in a hole is a tau transition not changing the action state.
\TODO{Id(s) is the identity function on variables of state s}
\begin{definition}[Non-observability of silent actions]
An open automaton $A = <J,\mathcal{S},s_0,\mathcal{T}>$ \emph{cannot observe silent actions} if and only if for all $j$ in $J$ and $s$ in $\mathcal{S}$ we have
\begin{enumerate}
\item
\[ \openrule
         {
           (j\mapsto\tau),\symb{True},\symb{Id}(s)}
         {s \OTarrow {\tau} s}
         \in \mathcal{T}
\]
and 
\item for all $\alpha$, $j$, $s$, $\Post$ and $s'$ such that
\[ \openrule
         {
           (j\mapsto\tau),\Pred,\Post}
         {s \OTarrow {\alpha} s'}
         \in \mathcal{T} \] we have \[ \alpha=\tau\land s=s'\land \Post=\symb{Id}(s)
\]
\end{enumerate}
\end{definition}

@@

\begin{property}[Weak open transition transitivity]
Let $A$ be an open automaton, and $<J,\mathcal{S},s_0,\WT>$ a weak open automaton derived from $A$.
Suppose we have the following weak open transitions in $\WT$:
\[\openrule
         {
           \overline{\gamma_1},\Pred_1,\Post_1   }
         {s \OTWeakarrow {\tau} s_1}
\qquad 
\openrule
         {
           \overline{\gamma_2},\Pred_2,\Post_2   }
         {s_1 \OTWeakarrow {\alpha} s_2}
\qquad
\openrule
         {
           \overline{\gamma_3},\Pred_3,\Post_3   }
         {s_2 \OTWeakarrow {\tau} s_3}
\]
Then there is a weak  open transition of the following form in $\WT$:
\[\openrule
         {
           \overline{\gamma},\Pred,\Post   }
         {s \OTWeakarrow {\alpha} s_3}
\]
Suppose additionally $\Post_3=({x_k\gets e_k})^{k\in K}$, we have the following
\[
\begin{array}{rcl}
\gamma&=& \gamma_1\dotcup\gamma_2\dotcup\gamma_3\\
\Pred&=&\Pred_1\land\Pred_2\subst{\Post_1}\land (\Pred_3\subst{\Post_2})\subst{\Post_1}\\
\Post &=& ({x_k\gets ((e_k\subst{\Post_2})\subst{\Post_1})})^{k\in K}
\end{array}\]
\end{property}

@@

Weak bisimulation can be alternatively defined as a strong bisimulation on the weak open automata, however this would still need a specific definition of strong bisimulation for weak open automata. \TODO{polish}

--------------------------------------------------------

We will show that $\mathcal{R}''$ is a weak FH-bisimulation. Consider 
$(s,u|\Pred_{s,u})\in\mathcal{R}''$. Then there is a set of states of $\mathcal{S}_3$ relating $s$ and $u$, let $(p)^{p\in \mathcal{S}_2}$ be this family.  We have $\Pred_{s,u} = \bigvee_{p\in \mathcal{S}_2} \Pred_{s,p}\land\Pred_{p,u}$.

For any $p\in \mathcal{S}_2$ by definition of $\mathcal{R}''$, $(s,p|\Pred_{s,p})\in\mathcal{R}$,  and $(p,u|\Pred_{p,u})\in\mathcal{R}'$. 
We have the following by definition of weak bisimulation:\\
For any open transition $OT$ in $\mathcal{T}_1$ originating from $s$.
       	\begin{mathpar}
       	\openrule
       	{
       		\beta_j^{j\in J_1},\Pred_{OT},\Post_{OT}}
       	{s \OTarrow {\alpha} {s}'}  \qquad (1)   	
       	\end{mathpar}

       	There exist weak open transitions $OT_{x}^{x\in X} \subseteq \mathcal{T}_2$:

       	\begin{mathpar} 
       	%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \
       	\openrule
       	{
       		\gamma_{j x}^{j\in J_{x}}, \Pred_{OT_{x}},\Post_{OT_{x}}}
       	{p \OTWeakarrow {\alpha_{x}} p'_{x}}\qquad (2)
       	%         \right)
       	\end{mathpar}

       	such that  $\forall x, J_1=J_{x}, (s',p'_{x}|\Pred\,'_{{x}})\in 
       	\mathcal{R}$;
       	and  \\

       	$\Pred_{s,p} \land \Pred_{OT}\\
       	\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       	\left( \forall j. \vis{\beta_j}=\gamma_{jx}  \Rightarrow \Pred_{OT_{x}}
       	\land \alpha\!=\!\alpha_{x} \land
       	\Pred\,'_{{x}}\subst{\Post_{OT}}\subst{\Post_{OT_{x}}}\right)$\\

For any open transition $OT_{x}$, since
       	$(p,u_x|\Pred_{x})\in\mathcal{R}'$ there exist open transitions
       	$OT_{xy}^{y\in Y} \subseteq \mathcal{T}_3$:

       	\begin{mathpar}  	
       	%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \
       	\openrule
       	{
       		\gamma_{j  x y}^{j\in J_{xy}}, 
       		\Pred_{OT_{xy}},\Post_{OT_{xy}}}
       	{u \OTWeakarrow {\alpha_{xy}} u'_{xy}}\qquad (3)
       	%         \right)    	
       	\end{mathpar}
       	such that  $\forall y, J_{x}=J_{xy}, 
       	(p'_{x},u'_{xy}|\Pred\,'_{{x,xy}})\in \mathcal{R}'$; and  \\
       	$\Pred_{p,u} \land \Pred_{OT_{x}}\\
       	\hspace{1cm} \implies\!\!\! \bigvee_{y\in Y}
       	\left( \forall j. \gamma_{jx}=\gamma_{jxy}  \Rightarrow \Pred_{OT_{xy}}
       	\land \alpha_{x}\!=\!\alpha_{xy} \land
       	\Pred\,'_{{x,xy}}\subst{\Post_{OT_{x}}}\subst{\Post_{OT_{xy}}}\right)$.\\

This is verified for all $p \in \mathcal{S}_2$. Overall,  we have a family of open 
       	transitions 
       	$OT_{xy}^{x\in X, 
       		y\in Y} \subseteq \mathcal{T}_3$ that should simulate \emph{OT}.

We have  $\forall y, \forall x,  J_1=J_{x}=J_{xy}, 
       	({s}',u'_{xy}|\Pred\,'_{{xy}})\in \mathcal{R}''$ for some $\Pred\,'_{{xy}}$.  Indeed,  for any $x$ and $y$, we have $p'_{x}$ relates ${s}'$ and $u'_{xy}$, indeed we have
$(s',p'_{x}|\Pred\,'_{{x}})\in \mathcal{R}$
       	and $(p'_{x},u'_{xy}|\Pred\,'_{{x,xy}})\in \mathcal{R}'$. For states ensuring the transitivity relation we have   
       		 for all  $x$, $y$, $\Pred\,'_{x}\land 
       	\Pred\,'_{{x,xy}}\implies 
       	\Pred\,'_{{xy}}$.

\TODO{We need to prove that we have the following  relation:}  \\     	
       	\begin{small}
 $\Pred_{s,u} \land \Pred_{OT}\\
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}\bigvee_{y\in Y}
       		\left( \forall j. \vis{\beta_j}=\gamma_{jxy}  \Rightarrow \Pred_{OT_{xy}}
       		\land \alpha\!=\!\alpha_{xy} \land
       		\Pred\,'_{{xy}}\subst{\Post_{OT}}\subst{\Post_{OT_{xy}}}\right)$.
       	\end{small}

\noindent By rule given in (2) we have for all state $p$:\\
\begin{small}
$\big(\Pred_{s,p} \land \Pred_{OT}\big)\land\Pred_{p,u}\\
       		\hspace{1cm} \implies\!\!\! \Big(\bigvee_{x\in X}
       		\left( \forall j. \vis{\beta_j}=\gamma_{jx}  \Rightarrow \Pred_{OT_{x}}
       		\land \alpha\!=\!\alpha_{x} \land
       		\Pred\,'_{{x}}\subst{\Post_{OT}}\subst{\Post_{OT_{x}}}\right)\Big)\land\Pred_{p,u}\\
       		%
       		\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       	\Big( \forall j. \vis{\beta_j}=\gamma_{jx}  \Rightarrow 
       		(\Pred_{OT_{x}}\land\Pred_{p,u})
       		\land \alpha\!=\!\alpha_{x} \land
       		\Pred\,'_{{x}}\subst{\Post_{OT}}\subst{\Post_{OT_{x}}}\Big).\\
       		%
       		$
  \end{small}

\noindent By rule given in (3) we get:     \\	
	\begin{small}
$    \big(\Pred_{s,p} \land \Pred_{OT}\big)\land\Pred_{p,u}\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
       		\bigg( \forall j. \vis{\beta_j}=\gamma_{jx}  \Rightarrow \Big(\bigvee_{y\in Y} 
       		\big( \forall j'.\gamma_{j'x}=\gamma_{j'xy}  \Rightarrow \Pred_{OT_{xy}}
       		\land \alpha_{x}\!=\!\alpha_{xy}\land\\
       		%
 \Pred\,'_{{x,xy}}\subst{\Post_{OT_{x}}}\subst{\Post_{OT_{xy}}}\Big)
       		\land \alpha\!=\!\alpha_{x} \land
       \Pred\,'_{{x}}\subst{\Post_{OT}}\subst{\Post_{OT_{x}}}\bigg)\\
       		%
 \hspace{1cm} \implies\!\!\! \bigvee_{x\in X} \bigvee_{y\in Y}
       		\bigg(\forall j, j'.\vis{\beta_j}=\gamma_{jx} \land \gamma_{j'x}=\gamma_{j'xy}
       		\Rightarrow \Big( 
       		\Pred_{OT_{xy}}
       		\land \alpha\!=\alpha_{x}\!=\!\alpha_{xy} \land
       \Pred\,'_{{x,xy}}\subst{\Post_{OT_{x}}}\subst{\Post_{OT_{xy}}}
       		\land
\Pred\,'_{{x}}\subst{\Post_{OT}}\subst{\Post_{OT_{x}}}\Big)\bigg).
       		$
       \end{small}

\medskip
\noindent Because of variables domains independence we  have:\\ 
    $\Pred\,'_{{x}}\subst{\Post_{OT}}\subst{\Post_{OT_{x}}}\land 
       	\Pred\,'_{{x,xy}}\subst{\Post_{OT_{x}}}\subst{\Post_{OT_{xy}}}
       	\Leftrightarrow\\ % we changed =
\Pred\,'_{{x}}\subst{\Post_{OT}}\subst{\Post_{OT_{x}}}\subst{\Post_{OT_{xy}}}\land
       \Pred\,'_{{x,xy}}\subst{\Post_{OT}}\subst{\Post_{OT_{x}}}\subst{\Post_{OT_{xy}}} \Leftrightarrow
\big(\Pred\,'_{{x}} \land   \Pred\,'_{{x,xy}}\big)   \subst{\Post_{OT}}\subst{\Post_{OT_{x}}}\subst{\Post_{OT_{xy}}} \\	\implies\Pred\,'_{{xy}}\subst{\Post_{OT}}\subst{\Post_{OT_{x}}}\subst{\Post_{OT_{xy}}}
   \Leftrightarrow  \Pred\,'_{{xy}}\subst{\Post_{OT}}\subst{\Post_{OT_{xy}}} $\\

\noindent This allows us to conclude, with $\Pred_{s,u} = \bigvee_{p\in \mathcal{S}_2} 
       	\Pred_{s,p}\land\Pred_{p,u}$:

      	\begin{small}     	
$\Pred_{s,u} \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \bigvee_{p\in \mathcal{S}_2} \big(
	\Pred_{s,p}\land\Pred_{p,u}\big) \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\!\bigvee_{p\in \mathcal{S}_2}\bigg(\!\!
 \bigvee_{x\in X}\!\!\bigvee_{y\in Y}\!\!
\Big( \forall j, j'.\vis{\beta_j}=\gamma_{jx} \land \gamma_{j'x}=\gamma_{j'xy}
\Rightarrow \Big( 
\Pred_{OT_{xy}}
\land \alpha\!=\alpha_{x}\!=\!\alpha_{xy}\\~\hspace{3em}~ \land
\Pred\,'_{{xy}}\subst{\Post_{OT}}\subst{\Post_{OT_{xy}}}\Big)\Big)\bigg)
\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}\!\!\bigvee_{y\in Y}\!\!\bigvee_{p\in \mathcal{S}_2}\!\!
\Big(\forall j.\vis{\beta_j}\!=\!\gamma_{jxy}  \Rightarrow \Pred_{OT_{xy}}
\land\, \alpha\!=\!\alpha_{xy} \land
\Pred\,'_{{xy}}\subst{\Post_{OT}}\subst{\Post_{OT_{xy}}}\Big)$.
  \end{small}

----------------------------------------------

@@

\subsection{Running example}
- Detail of a $\tau^*$ sequence, \\
- construction of the predicate\\
- defn / elimination of the intermediate variables\\
- construction of the weak OT.

The ``saturation'' step in the construction of the weak relation will
pick all $\tau^*$ and $\tau^*;a;\tau^*$ sequences from the OA, and build a weak
OT from each sequence. Here is one example of a short sequence,
starting with the open transition $OT_2$ of page \pageref{OT:ABP},
then transmitting the message to the receiver (if the sender and
receiver control bits are equal), and to the environment. The two
firts OTs are internal, the last one is visible, so it constitute a
$\tau;\tau;a$ weak transition.
